---
title: "Final Project Stat131A"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=60), tidy=TRUE)
library(tidyr)
library(tidyverse)
library(ggplot2)

```


```{r}
#read in dataset and change variables to factors
chol <- read.csv("cholangitis-data.csv")
chol[sapply(chol, is.character)] <- lapply(chol[sapply(chol, is.character)], as.factor)
chol$stage <-as.factor(chol$stage)
chol$id <- as.factor(chol$id)

#removing the NA cakuse and dividing age by 365 to make more legible 
chol1 <- chol %>% na.omit(cholangitis)
chol1 <- chol1 %>% mutate(age = age/365)
#str(chol)
```
We are removing all of the NA values that would skew our results. We are also dividing age by 365 to make the data a lot easier to understand. 

###Visualization
```{r}
#Boxplot with variables drug and n_days
drug.box<- chol1 %>% 
  ggplot(aes(drug, n_days)) + geom_boxplot()
drug.box
```


```{r}
#bar graph with that shows drug, placebo, and NA and the participants status 
status.bar<- chol1 %>% 
  ggplot(aes(status)) + geom_bar() +
  facet_grid(.~drug)
status.bar
```
It is interesting to note that with the D-penicillin drug and placebo drug, the results are fairly the same. There are almost the same number of alive and dead patients for both as well as those who have received a liver transplant. 

```{r}
age.box <- chol1 %>% 
  ggplot(aes(status, age)) + 
  geom_boxplot() +
  ylab('Age') + xlab('Status')
age.box
  
```
The key point to note in the boxplot isthat the younger aged demogrpahic is the ones who have received the liver transplant. 

```{r}
chol1 %>% 
  ggplot(aes(status)) + geom_bar() + facet_grid(.~sex) + theme_bw()

```
There are more female patients that survived compared to male though there 
was a lot more female patients to begin with. Because there are so few male
participants im not sure if sex will be a good explanatory variable to use.

```{r}
chol1 %>% 
ggplot(aes(x = drug, fill = status)) + geom_bar() + 
  xlab("Tyoe of Drug") + 
  ggtitle("Treatment") + 
  theme_bw()
```
```{r}
ggplot(chol1, aes(x = n_days, color = drug)) + geom_density() + 
  xlab("Number of Days") + 
  ggtitle("Days Enrolled Compared to Drug Treatment") + 
  geom_vline(xintercept = mean(filter(chol1, drug == 'D-penicillamine')$n_days), color = 'red', linetype="dotted") + 
  geom_vline(xintercept = mean(filter(chol1, drug == 'Placebo')$n_days), color = 'green', linetype="dotted") + theme_bw()
```

This graph shows the same data as the first boxplot but the lines help to
visualize the changes in the data over n_days while the boxplot is more useful 
for visualizing the means and IQR of the two drug groups.


```{r}
#Create a pairs plot of the continuous variables in the dataset
cont_var <- chol1[, -c(1, 3, 4, 6:10, 20)]
cat_var<- chol1[, c(3, 4, 6:10, 20)]
pairs(cont_var)
```

Multivariable Regression
```{r}
noCL <- chol %>% 
  filter(status != 'CL')
```
Filtering out the status of the patients to remove anybody who received a liver transplant because it cures the patients with primary biliary cholangitis. 


```{r}
quantile_limits <- function(df, lb, ub){
  lower_bound <- c()
  upper_bound <- c()
  for(i in 1:ncol(df)) {
    lower <- c(lower_bound, quantile(df[,i], lb))
    upper <- c(upper_bound, quantile(df[,i], ub))
  }
  output <- data.frame(names(df), lower, upper)
  return(output)
}

ci_cont <- quantile_limits(cont_var, 0.001, 0.999)
ci_cont

ci_95 <- quantile_limits(cont_var, 0.025, 0.975)


```
We are filtering our data by quantiles to prevent any unnecessary outliers from skewing the data.

```{r}
bil <- ggplot(chol1, aes(x = bilirubin, color = drug)) + geom_density() + ggtitle("Bilirubin Over Time") + theme(axis.text.x = element_text(angle = 45)) 

chl <- ggplot(chol1, aes(x = cholesterol, color = drug)) + geom_density() + ggtitle("Cholesterol Over Time") + theme(axis.text.x = element_text(angle = 45))

alb <- ggplot(chol1, aes(x = albumin, color = drug)) + geom_density() + ggtitle("Albumin Over Time") + theme(axis.text.x = element_text(angle = 45))

cop <- ggplot(chol1, aes(x = copper, color = drug)) + geom_density() + ggtitle("Copper Over Time") + theme(axis.text.x = element_text(angle = 45))

plot_grid(b, chl, a, cu)
```
```{r}
library(leaps)
no_id_refined_logs <- logs_only[, -c(1)]
cholregsub <- regsubsets(n_days ~., no_id_refined_logs, nvmax = 28) 
```

```{r}
cont_var_trans <- refined_df[, -c(1, 3, 4, 6:12, 14:17, 20)]
logs_only <- refined_df[, -c(11, 12, 14:17)]
full_lm <- lm(n_days ~ ., data = logs_only[, -c(1)])
summary(full_lm)
```

```{r}
summary(lm(formula = n_days ~ drug + ascites + edema + albumin + platelets + 
    stage + log_bilirubin + log_cholesterol + log_alk_phos + log_sgot, data = logs_only[, 
    -c(1)]))
```



graph drug treatments and survival by disease stage 

find out how to remove outliers from dataset first before grphing 

graph density plot for everything that is not factor compaing it to drug 

log all results and facet wrap it to see a better normal distribution 
